Paper 2 !DRAFT! Do not circulate!

### 2014-11-28 Articles/Article_EcolEcon_2014/Paper2.Rnw
Uncommented many lines for CES model.

% The CES model without and with energy is given by 
% Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
% The \texttt{R} \citep{R} package \texttt{micEconCES} 
% \citep{Henningsen:2011td} was used to estimate parameters 
% $\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$. 
% The \texttt{cesEst} function of \texttt{micEconCES} provides 
% several algorithm options for parameter estimation.
% The default algorithm (Levenberg-Marquardt) does not respect
% parameter constraints (see Section~\ref{sec:CES}) and, 
% in our testing, nearly always violated them, often returning
% negative values for elasticity of substitution~($\sigma$) parameters.
% Thus, we used the two fitting algorithms available in \texttt{cesEst} that
% respect coefficient constraints: \texttt{PORT} and \texttt{L-BFGS-B}. 
% Our CES parameter estimation algorithm starts with an eleven-value grid search
% in $\rho$ and $\rho_1$ (9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
% which corresponds to $\sigma$ and $\sigma_1$ values of
% 0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100, respectively.
% During the grid search, values of $\rho$ and $\rho_1$ are fixed,
% and values of $\lambda$, $\gamma$, $\delta$, and $\delta_1$ are estimated 
% by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms.
% In all, 121 gradient searches in $\lambda$, $\gamma$, $\delta$, and $\delta_1$
% at grid points representing all combinations of $\rho$ and $\rho_1$ are attempted.
% During the grid search portion of our algorithm, starting values for the free parameters
% are $\lambda$ = 0.015/year, $\delta$ = 0.5, $\delta_1$ = 0.5, 
% and $\gamma$ is set to a value such that the mean of the residuals is zero
% by the \texttt{cesEst} function.
% Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B}) 
% is attempted wherein all fitting parameters 
% ($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
% are allowed to float. 
% The start values for fitting parameters are taken from 
% the grid search fit that provided the lowest $sse$ (Equation~\ref{eq:sse}).
% If resampled data are being fitted, a prior fit to historical data is available.
% In the final step of our algorithm, 
% a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
% uses coefficients from the best 
% fit to historical data as its starting point. 
% In this final gradient search, all model parameters 
% ($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
% are considered free parameters.
% The fit with lowest $sse$ of all above trials is deemed the winner,
% and its fitting parameters are used as the model.
% \citet{Henningsen:2011td}, 
% in their detailed analysis of \citet{Kemfert:1998vd}, found that 
% \begin{quote}
% \ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
% are---at least in this study---most likely to find the 
% coefficients that give the best fit to the model, 
% where the \texttt{PORT} algorithm can be used to restrict the estimates 
% to the economically meaningful region.
% \end{quote}
% In our testing, we found that to be mostly true. 
% \texttt{PORT} nearly always provided lower
% $sse$ than \texttt{L-BFGS-B}, despite the fact that 
% \texttt{L-BFGS-B} often reports convergence
% and \texttt{PORT} does not for the same data
% (i.e., for the same $y$, $k$, $l$, and $e$ data).
% Coefficients for each resampled CES model for all countries and all nests
% are shown in 
% Figures~\ref{fig:CES_lambda_gamma_graph_X}--\ref{fig:CES_sigma_delta_graph_X}. 
% The white crosshair on each graph shows the location 
% of the fit to historical data. 
% In Figures~\ref{fig:CES_sigma1_delta1_graph_X} and~\ref{fig:CES_sigma_delta_graph_X},
% the bottom of the vertical axis represents $\sigma = 0$.
% The top of the vertical axis represents $\sigma = \infty$.
% Note that as $\delta$ or $\delta_1$ approaches 0 (1) 
% in Equations~\ref{eq:CES}--\ref{eq:CESekl},
% the corresponding value of $\sigma$ or $\sigma_1$ becomes irrelevant, and
% resample points along the left (right) edge of
% Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
% are equivalent.
% That is, all points along the left (right) edge of
% Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
% have the same $sse$. 
% As a consequence, 
% the \texttt{cesEst} fitting algorithm has no way 
% to discriminate among points along the left (right) edge of
% Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}. 
% Thus, resample point estimates where $\delta$ or $\delta_1$ approaches 0 (1)
% tend to cluster at the locations of the initial grid search in 
% $\sigma$ or $\sigma_1$
% (0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100).
% <<CES_lambda_gamma_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=4.5, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with exergy ($x$).", eval=FALSE>>=
% cesCoeffsDataX <- filter(coeffs_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
% standardScatterPlot(data=cesCoeffsDataX, mapping=aes(gamma_coef, lambda), 
%                     facet_formula=country~nestStrParen, alpha=0.05) +
% <<CES_sigma1_delta1_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=5, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with exergy ($x$).", eval=FALSE>>=
% standardScatterPlot(data=cesCoeffsDataX, mapping=aes(delta_1, sigma_1), facet_formula=country~nestStrParen, alpha=0.05) +
% <<CES_sigma_delta_graph_X, fig.pos="H", fig.align='center', fig.width=4.4, fig.height=5, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models using exergy ($x$).", eval=FALSE>>=
% standardScatterPlot(data=subset(cesCoeffsDataX, nestStrParen != "(iK + iL) + ()"),

>>>>

The CES model without and with energy is given by 
Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$. 
The \texttt{cesEst} function of \texttt{micEconCES} provides 
several algorithm options for parameter estimation.
The default algorithm (Levenberg-Marquardt) does not respect
parameter constraints (see Section~\ref{sec:CES}) and, 
in our testing, nearly always violated them, often returning
negative values for elasticity of substitution~($\sigma$) parameters.
Thus, we used the two fitting algorithms available in \texttt{cesEst} that
respect coefficient constraints: \texttt{PORT} and \texttt{L-BFGS-B}.

Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho$ and $\rho_1$ (9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
which corresponds to $\sigma$ and $\sigma_1$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100, respectively.
During the grid search, values of $\rho$ and $\rho_1$ are fixed,
and values of $\lambda$, $\gamma$, $\delta$, and $\delta_1$ are estimated 
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms.
In all, 121 gradient searches in $\lambda$, $\gamma$, $\delta$, and $\delta_1$
at grid points representing all combinations of $\rho$ and $\rho_1$ are attempted.
During the grid search portion of our algorithm, starting values for the free parameters
are $\lambda$ = 0.015/year, $\delta$ = 0.5, $\delta_1$ = 0.5, 
and $\gamma$ is set to a value such that the mean of the residuals is zero
by the \texttt{cesEst} function.

Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B}) 
is attempted wherein all fitting parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are allowed to float. 
The start values for fitting parameters are taken from 
the grid search fit that provided the lowest $sse$ (Equation~\ref{eq:sse}).

If resampled data are being fitted, a prior fit to historical data is available.
In the final step of our algorithm, 
a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
uses coefficients from the best 
fit to historical data as its starting point. 
In this final gradient search, all model parameters 
($\lambda$, $\gamma$, $\delta$, $\delta_1$, $\rho$, and $\rho_1$)
are considered free parameters.

The fit with lowest $sse$ of all above trials is deemed the winner,
and its fitting parameters are used as the model.

\citet{Henningsen:2011td}, 
in their detailed analysis of \citet{Kemfert:1998vd}, found that

\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the \texttt{PORT} algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ data).

Coefficients for each resampled CES model for all countries and all nests
are shown in 
Figures~\ref{fig:CES_lambda_gamma_graph_X}--\ref{fig:CES_sigma_delta_graph_X}. 
The white crosshair on each graph shows the location 
of the fit to historical data. 
In Figures~\ref{fig:CES_sigma1_delta1_graph_X} and~\ref{fig:CES_sigma_delta_graph_X},
the bottom of the vertical axis represents $\sigma = 0$.
The top of the vertical axis represents $\sigma = \infty$.

Note that as $\delta$ or $\delta_1$ approaches 0 (1) 
in Equations~\ref{eq:CES}--\ref{eq:CESekl},
the corresponding value of $\sigma$ or $\sigma_1$ becomes irrelevant, and
resample points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
are equivalent.
That is, all points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}
have the same $sse$. 
As a consequence, 
the \texttt{cesEst} fitting algorithm has no way 
to discriminate among points along the left (right) edge of
Figures \ref{fig:CES_sigma1_delta1_graph_X} and \ref{fig:CES_sigma_delta_graph_X}. 
Thus, resample point estimates where $\delta$ or $\delta_1$ approaches 0 (1)
tend to cluster at the locations of the initial grid search in 
$\sigma$ or $\sigma_1$
(0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100).

<<CES_lambda_gamma_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=4.5, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models with primary exergy ($x_p$).", eval=FALSE>>=
cesCoeffsDataX <- filter(AllCoeffs, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
                                                                      "(iK + iL) + (iXp)", 
                                                                      "(iL + iXp) + (iK)", 
                                                                      "(iXp + iK) + (iL)"))
standardScatterPlot(data=cesCoeffsDataX, mapping=aes(x=gamma_coef, y=lambda, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    labs(x=expression(gamma), y=expression(lambda)) + 
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<CES_sigma1_delta1_graph_X, fig.pos="H", fig.align='center', fig.width=5, fig.height=5, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models with primary exergy ($x_p$).", eval=FALSE>>=
standardScatterPlot(data=cesCoeffsDataX, mapping=aes(x=delta_1, y=sigma_1, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1])) + 
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@

<<CES_sigma_delta_graph_X, fig.pos="H", fig.align='center', fig.width=4.4, fig.height=5, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models using exergy ($x_p$).", eval=FALSE>>=
standardScatterPlot(data=subset(cesCoeffsDataX, nestStrParen != "(iK + iL) + ()"),
                    mapping=aes(x=delta, y=sigma, color=Source), 
                    facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma)) +
    guides(colour=guide_legend(override.aes = aes(alpha=1)))
@
